---
title: "Assignment 2: Coding Basics"
author: "Changxin Yu"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics Day 1

1.  Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r}
#1. 
seq0 <- seq(1,100,4) # generate a sequence

#2. 
mean(seq0) # compute the mean of the sequence
median(seq0) # compute the median of the sequence

#3. 
mean(seq0) > median(seq0) # determine if mean > median
```

## Basics Day 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r}
#5.
name <- c("Jack", "Anne", "Chris", "Sam") # character vector
score <- c(90, 67, 51, 46) # numeric vector
pass_or_not <- c(T,T,T,F) # logical vector

#6.
mode(name)
mode(score)
mode(pass_or_not)

#7.
students_grade <- as.data.frame(cbind(name,score,pass_or_not))

#8.
colnames(students_grade) # it already had informative column names
students_grade
```

9.  QUESTION: How is this data frame different from a matrix?

> Answer:For a matrix, all columns must have same mode and same length, but for a data frame, they should not be.

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement.

```{r}
pass_fun <- function(score){
  result <- ifelse(score<50, F, T)
  print(result)
}
```


11. Apply your function to the vector with test scores that you created in number 5.

```{r}
pass_fun(score)
```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: When we just do a simple logical judgement and one line of code for each case, ifelse works well. Otherwise, we should use if and else to fit more complicated cases.
