---
title: "Assignment 6: GLMs (Linear Regressions, ANOVA, & t-tests)"
author: "Changxin Yu"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#This code chunk will tidy your knit PDF files, wrapping long code lines
#For it to work, the "formatR" package needs to be installed

#install.packages('formatR')
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE,
                      warning=FALSE, message=FALSE)
```

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on generalized linear models. 

## Directions
1. Rename this file `<FirstLast>_A06_GLMs.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.


## Set up your session 
1. Set up your session. Check your working directory. Load the tidyverse, agricolae and other needed packages. Import the *raw* NTL-LTER raw data file for chemistry/physics (`NTL-LTER_Lake_ChemistryPhysics_Raw.csv`). Set date columns to date objects.

2. Build a ggplot theme and set it as your default theme.

```{r setup1}
#1
getwd()
setwd("E:/things/Duke University/study/2022 Fall/ENVIRON 872/EDA-Fall2022")
library(tidyverse)
library(agricolae)
library(lubridate)
NTL <- read.csv("./Data/Raw/NTL-LTER_Lake_ChemistryPhysics_Raw.csv",
                stringsAsFactors = TRUE)
NTL$sampledate <- as.Date(NTL$sampledate , format = "%m/%d/%y")

#2
mytheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

```

## Simple regression
Our first research question is: Does mean lake temperature recorded during July change with depth across all lakes?

3. State the null and alternative hypotheses for this question:

> Answer:
H0: The lake temperature recorded during July don't change with depth across all lakes, which means that the linear regression of July lake temperature and depth across all lakes has intercept and slope equaling zeros.
Ha: The lake temperature recorded during July do change with depth across all lakes. The linear regression of July lake temperature and depth across all lakes has nonzero intercept or slope.

4.  Wrangle your NTL-LTER dataset with a pipe function so that the records meet the following criteria: 
 * Only dates in July. 
 * Only the columns: `lakename`, `year4`, `daynum`, `depth`, `temperature_C`
 * Only complete cases (i.e., remove NAs)

5. Visualize the relationship among the two continuous variables with a scatter plot of temperature by depth. Add a smoothed line showing the linear model, and limit temperature values from 0 to 35 °C. Make this plot look pretty and easy to read.

```{r scatterplot}
#4
NTL1 <- NTL %>%
  mutate(Month = month(sampledate)) %>%
  filter(Month == 7) %>%
  select(lakename, year4, daynum, depth, temperature_C) %>%
  na.omit()

#5
ggplot(NTL1, aes(x = depth, y = temperature_C))+
  geom_point(size=0.8)+
  geom_smooth(method="lm")+
  ylim(0, 35)
```


6. Interpret the figure. What does it suggest with regards to the response of temperature to depth? Do the distribution of points suggest about anything about the linearity of this trend?

> Answer: The figure suggests that the lake temperature decreases when the depth increasing. The distribution of points shows that there is some of linearity in this trend, especially when depth is from 0 to 8 metres, but for the whole trend, it is more like an exponential relationship.


7. Perform a linear regression to test the relationship and display the results

```{r linear.regression}
#7
tempvsdepth <- lm(data=NTL1, temperature_C~depth)
summary(tempvsdepth)
```


8. Interpret your model results in words. Include how much of the variability in temperature is explained by changes in depth, the degrees of freedom on which this finding is based, and the statistical significance of the result. Also mention how much temperature is predicted to change for every 1m change in depth. 

> Answer: The R-squared of this linear model is 0.7387, which means that 73.87% variability in temperature is explained by changes in depth. The degrees of freedom is 9726 = 9728-2. The number of observations is 9728 and the number of parameters is 2. According to p-values of intercept and slope (both smaller than 0.05), the relationship between temperature and depth is significant so we can reject the null hypothesis. For every 1m increase in depth, the temperature decreases by 1.946 degrees according to the slope.

---

## Multiple regression
Let's tackle a similar question from a different approach. Here, we want to explore what might the best set of predictors for lake temperature in July across the monitoring period at the North Temperate Lakes LTER. 

9. Run an AIC to determine what set of explanatory variables (year4, daynum, depth) is best suited to predict temperature.

10. Run a multiple regression on the recommended set of variables. 

```{r temperature.model}
#9
TempAIC <- lm(data = NTL1, temperature_C ~ year4 + daynum + depth)
step(TempAIC)

#10
# The result of AIC shows that none of variables need to be removed.
summary(TempAIC)

```

11. What is the final set of explanatory variables that the AIC method suggests we use to predict temperature in our multiple regression? How much of the observed variance does this model explain? Is this an improvement over the model using only depth as the explanatory variable?

> Answer: The final set of explanatory variables according to AIC result is "year4 + daynum + depth". The new model explains 74.12% variance, which is a bit greater than 73.87% from the previous univariate model, so there is an improvement in new model. 


---
## Analysis of Variance

12. Now we want to see whether the different lakes have, on average, different temperatures in the month of July. Run an ANOVA test to complete this analysis. (No need to test assumptions of normality or similar variances.) Create two sets of models: one expressed as an ANOVA model and another expressed as a linear model (as done in our lessons).

```{r anova.model}
#12
# ANOVA model
lake.anova1 <- aov(data = NTL1, temperature_C ~ lakename)
summary(lake.anova1)

# linear model
lake.anova2 <- lm(data = NTL1, temperature_C ~ lakename)
summary(lake.anova2)

```

13. Is there a significant difference in mean temperature among the lakes? Report your findings. 

> Answer: From the ANOVA test, the p-value is smaller than 0.05, so there is a significant difference in mean temperature among the lakes.


14. Create a graph that depicts temperature by depth, with a separate color for each lake. Add a geom_smooth (method = "lm", se = FALSE) for each lake. Make your points 50 % transparent. Adjust your y axis limits to go from 0 to 35 degrees. Clean up your graph to make it pretty. 

```{r scatterplot.2}
#14.
ggplot(NTL1, aes(x = depth, y = temperature_C, color = lakename))+
  geom_point(size=0.5, alpha=0.5)+
  geom_smooth(method="lm", se = FALSE, size=0.8)+
  ylim(0, 35)+
  theme(legend.text = element_text(size = 9), 
        legend.title = element_text(size=11))+
  scale_color_brewer(palette = "PuBuGn")

```

15. Use the Tukey's HSD test to determine which lakes have different means.

```{r tukey.test}
#15
TukeyHSD(lake.anova1)
lake.groups <- HSD.test(lake.anova1, 'lakename', group = TRUE)
lake.groups
```

16.From the findings above, which lakes have the same mean temperature, statistically speaking, as Peter Lake? Does any lake have a mean temperature that is statistically distinct from all the other lakes?

>Answer: The Paul Lake has the same mean temperature as Peter Lake statistically with difference equaling -0.498 and p-value equaling 0.224. According to p-values, there is no lake having a mean temperature that is statistically distinct from all the other lakes.

 
17. If we were just looking at Peter Lake and Paul Lake. What's another test we might explore to see whether they have distinct mean temperatures? 

>Answer: T-test. The null hypothesis is that the mean temperature is the same between Peter Lake and Paul Lake.


18. Wrangle the July data to include only records for Crampton Lake and Ward Lake. Run the two-sample T-test on these data to determine whether their July temperature are same or different. What does the test say? Are the mean temperatures for the lakes equal? Does that match you answer for part 16?

```{r t.test}
NTL2 <- filter(NTL1, lakename == 'Crampton Lake' | lakename == 'Ward Lake')
lake.ttest <- t.test(NTL2$temperature_C ~ NTL2$lakename)
lake.ttest
```

>Answer: The test suggests that the mean temperature of these two lakes are euqal with p-value equaling 0.2649. That matches the answer from part 16.
